### 2015-05-04 internal_review_response.Rnw

Page 7: setstep.swmpr??

{\it See response to the next comment.}

Page 7: Not sure how best to refer to these S3 methods. You do it both ways (with .smpr in the table) and without in the text. Could keep it that way if you add a line to table caption. I bring it up, because I think new R users would be confused by this (I was for awhile...)

{\it Agreed, I've removed the method identifier from each function in the tables to avoid confusion.}

@@

Page 9: Does PLOS allow Boxes? Like Box 3. Example of qaqc() funciton usage... That would be a good way to show these.

{\it I've checked a few papers from PLOS One that describe software packages and most follow the format I've used, i.e., presentation of code preceded by the command prompt.}

@@

Page 9: quote

{\it Added.}

Page 10: A section on package dependencies might be a good addition.

{\it I've added a footnote that directs readers to the package index page on CRAN where the dependencies and package imports can be viewed.  I'm not sure it's worth describing each in detail since this is readily available in the source code.}

### 2015-05-04 swmpr_manu.Rnw
through the part about package dependencies

funcs <- c('comb.swmpr', 'qaqc.swmpr', 'qaqcchk.swmpr', 'rem\\_reps.swmpr', 'setstep.swmpr', 'subset.swmpr')
funcs <- paste0('\\texttt{', funcs, '}')

descrips <- c(
  'Combines \\texttt{swmpr} objects to a common time series using setstep, such as combining the weather, nutrients, and water quality data for a single station. Only different data types can be combined.',
  'Remove \\ac{QAQC} columns and remove data based on \\ac{QAQC} flag values for a \\texttt{swmpr} object.  Only applies if \\ac{QAQC} columns are present. ', 
  'View a summary of the number of observations in a \\texttt{swmpr} object that are assigned to different \\ac{QAQC} flags used by \\ac{CDMO}.  The output can be used to inform further processing.', 
  'Remove replicate nutrient data that occur on the same day.  The default is to average replicates.', 
  'Format data from a \\texttt{swmpr} object to a continuous time series at a given timestep.  The function is used in \\texttt{comb.swmpr} and can also be used with individual stations.',
  'Subset by dates and/or columns for a \\texttt{swmpr} object.  This is a method passed to the generic \\texttt{subset} function provided in the base package.'
  )

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = "Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?comb.swmpr} at the command line).",
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:organize', 
  table.env = FALSE
  )
@

@@

Raw nutrient data obtained from the \ac{CDMO} will usually include replicate samples that were taken within a few minutes of each other.  The \texttt{rem\_reps.swmpr} function combines nutrient data that occur on the same day to preserve an approximate monthly time step.  The \texttt{datetimestamp} column will always be averaged for replicates, but the actual observations will be combined based on the user-supplied function which defaults to the mean.  Other suggested functions include the \texttt{median}, \texttt{min}, or \texttt{max}.  The entire function call, including treatment of \texttt{NA}, values should be passed to the \texttt{FUN} argument (see the examples).  The function is meant to be used after \texttt{qaqc} processing, although it works with a warning if \ac{QAQC} columns are present.

@@

The \texttt{comb} function is used to combine multiple \texttt{swmpr} objects into a single object with a continuous time series at a given step.  The \texttt{timestep} function is used internally such that \texttt{timestep} and \texttt{differ} are accepted arguments for \texttt{comb}.  All arguments must be called explicitly since an arbitrary number of \texttt{swmpr} objects can be used as input.  The function combines data by creating a master time series that is used to iteratively merge all \texttt{swmpr} objects.  The time series for merging depends on the value passed to the \texttt{method} argument.  Passing \texttt{union} to \texttt{method} will create a time series that is continuous from the earliest and latest dates for all input objects.  Passing \texttt{intersect} to \texttt{method} will create a time series that is continuous from the set of dates that are shared between all input objects.  Finally, a seven or eight character station name passed to \texttt{method} will merge all data based on a continuous time series for the specified station, which must be present in the input data.  Currently, combining identical data types from different stations is not possible (e.g., two water quality stations from the same reserve).

@@

The analysis functions range from general purpose tools for time series analysis to more specific functions for working with continuous monitoring data in estuaries (\Cref{tab:analyze}).  The general purpose tools are \texttt{swmpr} methods that were developed for existing generic functions in the R base installation or relevant packages.  These functions include \texttt{swmpr} methods for \texttt{aggreswmp}, \texttt{filter}, and \texttt{approx} to deal with missing or noisy data and more general functions for exploratory data analysis, such as \texttt{plot}, \texttt{lines}, and \texttt{hist} methods.  Decomposition functions, \texttt{decomp} and \texttt{decomp\_cj}, are provided as relatively simple approaches for decomposing time series into additive or multiplicative components. Functions to estimate and plot ecosystem metabolism from combined water quality and weather data are provided by the \texttt{ecometab} and \texttt{plot\_metab} functions.  The analysis functions may or may not return a \texttt{swmpr} object depending on whether further processing with \texttt{swmpr} methods is possible from the output.

@@

funcs <- c('aggreswmp.swmpr', 'aggremetab.swmpr', 'ecometab.swmpr', 'decomp.swmpr', 'decomp\\_cj.swmpr', 'hist.swmpr', 'lines.swmpr', 'na.approx.swmpr', 'plot.swmpr', 'plot\\_metab.swmpr', 'plot\\_summary.swmpr', 'smoother.swmpr')
funcs <- paste0('\\texttt{', funcs, '}')

descrips <- c(
  'Aggregate \\texttt{swmpr} objects for different time periods - years, quarters, months,  weeks, days, or hours.  The aggregation function defaults to the mean.',
  'Aggregate metabolism data from a \\texttt{swmpr} object.  This is primarly used within \\texttt{plot\\_metab} but may be useful for simple summaries of raw metabolism data.',
  'Estimate ecosystem metabolism for a combined water quality and weather dataset using the open-water method.',
  'Decompose a \\texttt{swmpr} time series into trend, seasonal, and residual components.  This is a simple wrapper to \\texttt{decompose} \\cite{Kendall83}.  Decomposition of monthly or daily trends is possible.',
  'Decompose a \\texttt{swmpr} time series into grandmean, annual, seasonal, and events components.  This is a simple wrapper to \\texttt{decompTs} in the wq package \\cite{Jassby14}.  Only monthly decomposition is possible.',
  'Plot a histogram for a \\texttt{swmpr} object.',
  'Add lines to an existing plot created with \\texttt{plot.swmpr}.',
  'Linearly interpolate missing data (\\texttt{NA} values) in a \\texttt{swmpr} object. The maximum gap size that is interpolated is defined by the arguments.', 
  'Plot a univariate  time series for a \\texttt{swmpr} object.  The parameter name must be specified.',
  'Plot ecosystem metabolism estimates after running \\texttt{ecometab} on a \\texttt{swmpr} object.',  
  'Create summary plots of seasonal/annual trends and anomalies for a water a single paramter of interest.',
  'Smooth \\texttt{swmpr} objects with a moving window average.  Window size and sides (e.g., centered) can be specified, passed to \\texttt{filter}.'
  )

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggreswmp.swmpr} at the command line).",
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:analyze', 
  table.env = FALSE
  )
@

>>>>

funcs <- c('comb', 'qaqc', 'qaqcchk', 'rem\\_reps', 'setstep', 'subset')
funcs <- paste0('\\texttt{', funcs, '}')

descrips <- c(
  'Combines \\texttt{swmpr} objects to a common time series using setstep, such as combining the weather, nutrients, and water quality data for a single station. Only different data types can be combined.',
  'Remove \\ac{QAQC} columns and remove data based on \\ac{QAQC} flag values for a \\texttt{swmpr} object.  Only applies if \\ac{QAQC} columns are present. ', 
  'View a summary of the number of observations in a \\texttt{swmpr} object that are assigned to different \\ac{QAQC} flags used by \\ac{CDMO}.  The output can be used to inform further processing.', 
  'Remove replicate nutrient data that occur on the same day.  The default is to average replicates.', 
  'Format data from a \\texttt{swmpr} object to a continuous time series at a given timestep.  The function is used in \\texttt{comb} and can also be used with individual stations.',
  'Subset by dates and/or columns for a \\texttt{swmpr} object.  This is a method passed to the generic \\texttt{subset} function provided in the base package.'
  )

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = "Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?comb} at the command line).",
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:organize', 
  table.env = FALSE
  )
@

@@

Raw nutrient data obtained from the \ac{CDMO} will usually include replicate samples that were taken within a few minutes of each other.  The \texttt{rem\_reps} function combines nutrient data that occur on the same day to preserve an approximate monthly time step.  The \texttt{datetimestamp} column will always be averaged for replicates, but the actual observations will be combined based on the user-supplied function which defaults to the mean.  Other suggested functions include the \texttt{median}, \texttt{min}, or \texttt{max}.  The entire function call, including treatment of \texttt{NA}, values should be passed to the \texttt{FUN} argument (see the examples).  The function is meant to be used after \texttt{qaqc} processing, although it works with a warning if \ac{QAQC} columns are present.

@@

The \texttt{comb} function is used to combine multiple \texttt{swmpr} objects into a single object with a continuous time series at a given step.  The \texttt{timestep} function is used internally such that \texttt{timestep} and \texttt{differ} are accepted arguments for \texttt{comb}.  All arguments must be called explicitly since an arbitrary number of \texttt{swmpr} objects can be used as input.  The function combines data by creating a master time series that is used to iteratively merge all \texttt{swmpr} objects.  The time series for merging depends on the value passed to the \texttt{method} argument.  Passing \texttt{`union'} to \texttt{method} will create a time series that is continuous from the earliest and latest dates for all input objects.  Passing \texttt{`intersect`} to \texttt{method} will create a time series that is continuous from the set of dates that are shared between all input objects.  Finally, a seven or eight character station name passed to \texttt{method} will merge all data based on a continuous time series for the specified station, which must be present in the input data.  Currently, combining identical data types from different stations is not possible (e.g., two water quality stations from the same reserve).

@@

The analysis functions range from general purpose tools for time series analysis to more specific functions for working with continuous monitoring data in estuaries (\Cref{tab:analyze}).  The general purpose tools are \texttt{swmpr} methods that were developed for existing generic functions in the R base installation or relevant packages\footnote{See the listing on \ac{CRAN} for package imports and dependencies for SWMPr: \url{http://cran.r-project.org/web/packages/SWMPr/index.html}}.  These functions include \texttt{swmpr} methods for \texttt{aggreswmp}, \texttt{filter}, and \texttt{approx} to deal with missing or noisy data and more general functions for exploratory data analysis, such as \texttt{plot}, \texttt{lines}, and \texttt{hist} methods.  Decomposition functions, \texttt{decomp} and \texttt{decomp\_cj}, are provided as relatively simple approaches for decomposing time series into additive or multiplicative components. Functions to estimate and plot ecosystem metabolism from combined water quality and weather data are provided by the \texttt{ecometab} and \texttt{plot\_metab} functions.  The analysis functions may or may not return a \texttt{swmpr} object depending on whether further processing with \texttt{swmpr} methods is possible from the output.

@@

funcs <- c('aggreswmp', 'aggremetab', 'ecometab', 'decomp', 'decomp\\_cj', 'hist', 'lines', 'na.approx', 'plot', 'plot\\_metab', 'plot\\_summary', 'smoother')
funcs <- paste0('\\texttt{', funcs, '}')

descrips <- c(
  'Aggregate \\texttt{swmpr} objects for different time periods - years, quarters, months,  weeks, days, or hours.  The aggregation function defaults to the mean.',
  'Aggregate metabolism data from a \\texttt{swmpr} object.  This is primarly used within \\texttt{plot\\_metab} but may be useful for simple summaries of raw metabolism data.',
  'Estimate ecosystem metabolism for a combined water quality and weather dataset using the open-water method.',
  'Decompose a \\texttt{swmpr} time series into trend, seasonal, and residual components.  This is a simple wrapper to \\texttt{decompose} \\cite{Kendall83}.  Decomposition of monthly or daily trends is possible.',
  'Decompose a \\texttt{swmpr} time series into grandmean, annual, seasonal, and events components.  This is a simple wrapper to \\texttt{decompTs} in the wq package \\cite{Jassby14}.  Only monthly decomposition is possible.',
  'Plot a histogram for a \\texttt{swmpr} object.',
  'Add lines to an existing plot created with \\texttt{plot}.',
  'Linearly interpolate missing data (\\texttt{NA} values) in a \\texttt{swmpr} object. The maximum gap size that is interpolated is defined by the arguments.', 
  'Plot a univariate  time series for a \\texttt{swmpr} object.  The parameter name must be specified.',
  'Plot ecosystem metabolism estimates after running \\texttt{ecometab} on a \\texttt{swmpr} object.',  
  'Create summary plots of seasonal/annual trends and anomalies for a water a single paramter of interest.',
  'Smooth \\texttt{swmpr} objects with a moving window average.  Window size and sides (e.g., centered) can be specified, passed to \\texttt{filter}.'
  )

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggreswmp} at the command line).",
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:analyze', 
  table.env = FALSE
  )
@

### 2015-05-04 swmpr_manu.tex
through the part about package dependencies

The attributes for a \texttt{swmpr} object are descriptors that are appended to the raw data (\cref{tab:attributes}).  These act as metadata that are used internally by many of the package functions and are updated as the data are processed. Attributes of a \texttt{swmpr} object can also be viewed in R as follows.

@@

% table for organize functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?comb.swmpr} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:organize",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?comb.swmpr} at the command line).\label{tab:organize}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{comb.swmpr}&Combines \texttt{swmpr} objects to a common time series using setstep, such as combining the weather, nutrients, and water quality data for a single station. Only different data types can be combined.\tabularnewline
\texttt{qaqc.swmpr}&Remove \ac{QAQC} columns and remove data based on \ac{QAQC} flag values for a \texttt{swmpr} object.  Only applies if \ac{QAQC} columns are present. \tabularnewline
\texttt{qaqcchk.swmpr}&View a summary of the number of observations in a \texttt{swmpr} object that are assigned to different \ac{QAQC} flags used by \ac{CDMO}.  The output can be used to inform further processing.\tabularnewline
\texttt{rem\_reps.swmpr}&Remove replicate nutrient data that occur on the same day.  The default is to average replicates.\tabularnewline
\texttt{setstep.swmpr}&Format data from a \texttt{swmpr} object to a continuous time series at a given timestep.  The function is used in \texttt{comb.swmpr} and can also be used with individual stations.\tabularnewline
\texttt{subset.swmpr}&Subset by dates and/or columns for a \texttt{swmpr} object.  This is a method passed to the generic \texttt{subset} function provided in the base package.\tabularnewline
\hline
\end{tabular}\end{center}

@@

% table for analysis functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggreswmp.swmpr} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:analyze",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \texttt{?aggreswmp.swmpr} at the command line).\label{tab:analyze}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{aggreswmp.swmpr}&Aggregate \texttt{swmpr} objects for different time periods - years, quarters, months,  weeks, days, or hours.  The aggregation function defaults to the mean.\tabularnewline
\texttt{aggremetab.swmpr}&Aggregate metabolism data from a \texttt{swmpr} object.  This is primarly used within \texttt{plot\_metab} but may be useful for simple summaries of raw metabolism data.\tabularnewline
\texttt{ecometab.swmpr}&Estimate ecosystem metabolism for a combined water quality and weather dataset using the open-water method.\tabularnewline
\texttt{decomp.swmpr}&Decompose a \texttt{swmpr} time series into trend, seasonal, and residual components.  This is a simple wrapper to \texttt{decompose} \cite{Kendall83}.  Decomposition of monthly or daily trends is possible.\tabularnewline
\texttt{decomp\_cj.swmpr}&Decompose a \texttt{swmpr} time series into grandmean, annual, seasonal, and events components.  This is a simple wrapper to \texttt{decompTs} in the wq package \cite{Jassby14}.  Only monthly decomposition is possible.\tabularnewline
\texttt{hist.swmpr}&Plot a histogram for a \texttt{swmpr} object.\tabularnewline
\texttt{lines.swmpr}&Add lines to an existing plot created with \texttt{plot.swmpr}.\tabularnewline
\texttt{na.approx.swmpr}&Linearly interpolate missing data (\texttt{NA} values) in a \texttt{swmpr} object. The maximum gap size that is interpolated is defined by the arguments.\tabularnewline
\texttt{plot.swmpr}&Plot a univariate  time series for a \texttt{swmpr} object.  The parameter name must be specified.\tabularnewline
\texttt{plot\_metab.swmpr}&Plot ecosystem metabolism estimates after running \texttt{ecometab} on a \texttt{swmpr} object.\tabularnewline
\texttt{plot\_summary.swmpr}&Create summary plots of seasonal/annual trends and anomalies for a water a single paramter of interest.\tabularnewline
\texttt{smoother.swmpr}&Smooth \texttt{swmpr} objects with a moving window average.  Window size and sides (e.g., centered) can be specified, passed to \texttt{filter}.\tabularnewline
\hline
\end{tabular}\end{center}

>>>>

The attributes for a \texttt{swmpr} object are descriptors that are appended to the raw data (\cref{tab:attributes}).  These act as metadata that are used internally by many of the package functions and are updated as the data are processed. The attributes are not visible with the raw data but can be viewed as follows.

@@

% table for organize functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?comb} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:organize",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Organizing functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?comb} at the command line).\label{tab:organize}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{comb}&Combines \texttt{swmpr} objects to a common time series using setstep, such as combining the weather, nutrients, and water quality data for a single station. Only different data types can be combined.\tabularnewline
\texttt{qaqc}&Remove \ac{QAQC} columns and remove data based on \ac{QAQC} flag values for a \texttt{swmpr} object.  Only applies if \ac{QAQC} columns are present. \tabularnewline
\texttt{qaqcchk}&View a summary of the number of observations in a \texttt{swmpr} object that are assigned to different \ac{QAQC} flags used by \ac{CDMO}.  The output can be used to inform further processing.\tabularnewline
\texttt{rem\_reps}&Remove replicate nutrient data that occur on the same day.  The default is to average replicates.\tabularnewline
\texttt{setstep}&Format data from a \texttt{swmpr} object to a continuous time series at a given timestep.  The function is used in \texttt{comb} and can also be used with individual stations.\tabularnewline
\texttt{subset}&Subset by dates and/or columns for a \texttt{swmpr} object.  This is a method passed to the generic \texttt{subset} function provided in the base package.\tabularnewline
\hline
\end{tabular}\end{center}

@@

% table for analysis functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggreswmp} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:analyze",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \texttt{?aggreswmp} at the command line).\label{tab:analyze}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{aggreswmp}&Aggregate \texttt{swmpr} objects for different time periods - years, quarters, months,  weeks, days, or hours.  The aggregation function defaults to the mean.\tabularnewline
\texttt{aggremetab}&Aggregate metabolism data from a \texttt{swmpr} object.  This is primarly used within \texttt{plot\_metab} but may be useful for simple summaries of raw metabolism data.\tabularnewline
\texttt{ecometab}&Estimate ecosystem metabolism for a combined water quality and weather dataset using the open-water method.\tabularnewline
\texttt{decomp}&Decompose a \texttt{swmpr} time series into trend, seasonal, and residual components.  This is a simple wrapper to \texttt{decompose} \cite{Kendall83}.  Decomposition of monthly or daily trends is possible.\tabularnewline
\texttt{decomp\_cj}&Decompose a \texttt{swmpr} time series into grandmean, annual, seasonal, and events components.  This is a simple wrapper to \texttt{decompTs} in the wq package \cite{Jassby14}.  Only monthly decomposition is possible.\tabularnewline
\texttt{hist}&Plot a histogram for a \texttt{swmpr} object.\tabularnewline
\texttt{lines}&Add lines to an existing plot created with \texttt{plot}.\tabularnewline
\texttt{na.approx}&Linearly interpolate missing data (\texttt{NA} values) in a \texttt{swmpr} object. The maximum gap size that is interpolated is defined by the arguments.\tabularnewline
\texttt{plot}&Plot a univariate  time series for a \texttt{swmpr} object.  The parameter name must be specified.\tabularnewline
\texttt{plot\_metab}&Plot ecosystem metabolism estimates after running \texttt{ecometab} on a \texttt{swmpr} object.\tabularnewline
\texttt{plot\_summary}&Create summary plots of seasonal/annual trends and anomalies for a water a single paramter of interest.\tabularnewline
\texttt{smoother}&Smooth \texttt{swmpr} objects with a moving window average.  Window size and sides (e.g., centered) can be specified, passed to \texttt{filter}.\tabularnewline
\hline
\end{tabular}\end{center}
